50sites

### 2020-10-20 50sites files/50sites.Rmd
updated prior analysis for figures

  # here we use all patient samples, excluding glands
  #indices <- c(10:14, 16:40,47,48,58,59,73:76,87:96) # if using full data file
  indices <- c(10:57) # if using bulk only data file
  patientLabel <- substr(colnames(temp[indices]),1,1)
  patientLabel[10:12] <- "K*"
  sideLabel <- substr(colnames(temp[indices]),2,2)
  tissueLabel <- sideLabel
  tissueLabel[!(sideLabel %in% c("N"))] <- "T"   #replace A/B/D/M with T
  tumorIndicator <- 1*(tissueLabel=="T")

  work <- data.frame(beta=logit(t(temp[indices])), patient = patientLabel, tissue=tissueLabel, side=sideLabel, tInd=tumorIndicator)
  colnames(work)[1] <- "beta"
  return(work)
}

# Using Model with TCGA priors (previously model 3)
stanfit_model <- function (dataset) {
  stanDat <- list(
    N = nrow(dataset),
    P = nlevels(dataset$patient),
    pID = as.integer(factor(dataset$patient)),
    tInd = dataset$tInd,
    y = dataset[, 1]
    )
  stanFit <-
    stan(
      fit = emptyFit,
      warmup = 200,
      iter = 2000,
      chains = 4,
      data = stanDat,
      control = list(adapt_delta = 0.999),
      refresh = 0,
      seed=123
    )
  return(stanFit=stanFit)
}

@@

# Plot posterior distribution
postplot <- function(stanFit) {
  p2 <- mcmc_areas(
    as.array(stanFit), 
    pars = c("sigma_p", "sigma_pt","sigma_t","sigma_e"),
    prob = 0.8, # 80% intervals
    prob_outer = 0.95, 
    point_est = "median"
  )
  return(p2)
}

@@

# check for model_TCGApriors.rds, must add to prevent crash
checkFile <- "model_relaxedgamma5.rds"
if (file.exists(checkFile)) {
  file.remove(checkFile)
  print("caught one!") 
}

# create empty fit to use same model compilation throughout
data1<-site(1)
stanDat1 <- list(
  N = nrow(data1),
  P = nlevels(data1$patient),
  pID = as.integer(factor(data1$patient)),
  tInd = data1$tInd,
  y = data1[, 1]
)
emptyFit <- stan(file="model_relaxedgamma5.stan", data = stanDat1, chains = 0)

```{r echo=FALSE,include=TRUE,error=FALSE,message=FALSE,warning=FALSE,results='asis',fig.keep='all',fig.show='asis',fig.height=2.9,fig.width=5}
# fig.align='center'
ptm <- proc.time()
n_eff_sum<-numeric(6)
for (i in siteInds) {
  # fit model at given site
  data <- site(i)
  stanFit <- stanfit_model(data)

  # plot the fit plot (data and fit overlays), and posteriors of sigmas
  p1 <- fitplot(data, stanFit) + ggtitle(paste("Site", i, ": Beta Fits by Patient"))
  p2 <- postplot(stanFit) + ggtitle("Sigma Posteriors")
  p3 <- traceplot(stanFit,pars=c("sigma_p","sigma_pt","sigma_t","sigma_e"),inc_warmup=TRUE,ncol=1,window=c(0,1000)) + theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_text(angle=45),legend.position="none")
  p4 <- ggarrange(p2,p3,nrow=1,ncol=2)
  print(p1)
  print(p4)

  # text output:
  cat(paste("  \nSite ",i,": Parameter posterior summary\n",sep=''))
  # print (or table) mean, median, R_hat, n_eff for each of the 6 params
  datasumm <-summary(stanFit)$summary[1:6,c(1,6,9,10)]
  print(kable(datasumm))
  cat("  \n")
  n_eff_sum <- n_eff_sum + datasumm[,3]
}
mtp <- proc.time() - ptm

>>>>

  # here we use all patient samples, excluding glands
  #indices <- c(10:14, 16:40,47,48,58,59,73:76,87:96) # if using full data file
  indices <- c(10:57) # if using bulk only data file
  patientLabel <- substr(colnames(temp[indices]), 1, 1)
  patientLabel[10:12] <- "K*"
  sideLabel <- substr(colnames(temp[indices]), 2, 2)
  tissueLabel <- sideLabel
  tissueLabel[!(sideLabel %in% c("N"))] <- "T"   #replace A/B/D/M with T
  tumorIndicator <- 1 * (tissueLabel=="T")

  work <- data.frame(
    beta = logit(t(temp[indices])),
    patient = patientLabel,
    tissue = tissueLabel,
    side = sideLabel,
    tInd = tumorIndicator
    )
  colnames(work)[1] <- "beta"
  return(work)
}

# Using Model with TCGA priors (previously model 3)
stanfit_model <- function (dataset) {
  stanDat <- list(
    pID = as.integer(factor(dataset$patient)),
    tInd = dataset$tInd,
    N = nrow(dataset),
    P = nlevels(dataset$patient),
    y = dataset[, 1]
    )
  stanFit <-
    stan(
      fit = emptyFit,
      warmup = 200,
      iter = 2000,
      chains = 4,
      data = stanDat,
      control = list(adapt_delta = 0.999),
      refresh = 0,
      seed=123
    )
  return(stanFit=stanFit)
}

@@

# Plot posterior distribution
postplot_mubetaT <- function(stanFit) {
  p2 <- mcmc_areas(
    as.array(stanFit), 
    pars = c("mu","betaT"),
    prob = 0.8, # 80% intervals
    prob_outer = 0.95, 
    point_est = "median"
  )
  return(p2)
}
postplot_sigmas <- function(stanFit) {
  p2 <- mcmc_areas(
    as.array(stanFit), 
    pars = c("sigma_p", "sigma_pt","sigma_t","sigma_e"),
    prob = 0.8, # 80% intervals
    prob_outer = 0.95, 
    point_est = "median"
  )
  return(p2)
}

@@

# check for model_TCGApriors.rds, must add to prevent crash
checkFile <- "model_relaxedgamma3.rds"
if (file.exists(checkFile)) {
  file.remove(checkFile)
  print("caught one!") 
}

# create empty fit to use same model compilation throughout
data1<-site(1)
stanDat1 <- list(
  N = nrow(data1),
  P = nlevels(data1$patient),
  pID = as.integer(factor(data1$patient)),
  tInd = data1$tInd,
  y = data1[, 1]
)
emptyFit <- stan(file="model_relaxedgamma3.stan", data = stanDat1, chains = 0)

# inds for extracting data (mean, sem, median, n_eff, Rhat)
mInd <- c(1, 2, 6, 9, 10)
```

```{r echo=FALSE,include=TRUE,error=FALSE,message=FALSE,warning=FALSE,results='asis',fig.keep='all',fig.show='asis',fig.height=2.9,fig.width=5}

ptm <- proc.time()
n_eff_sum<-numeric(6)
for (i in siteInds) {
  # fit model at given site
  data <- site(i)
  stanFit <- stanfit_model(data)

  # plot the fit plot (data and fit overlays), and posteriors of sigmas
  p1 <- fitplot(data, stanFit) + ggtitle(paste("Site", i, ": Beta Fits by Patient"))
  p15 <- postplot_mubetaT(stanFit) + ggtitle("Mu/BetaT Posteriors")
  p2 <- postplot_sigmas(stanFit) + ggtitle("Sigma Posteriors")
  p3 <- traceplot(stanFit,pars=c("sigma_p","sigma_pt","sigma_t","sigma_e"),inc_warmup=TRUE,ncol=1,window=c(0,1000)) + theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_text(angle=45),legend.position="none")
  p4 <- ggarrange(p2,p3,nrow=1,ncol=2)
  print(p1)
  print(p4)

  # text output:
  cat(paste("  \nSite ",i,": Parameter posterior summary\n",sep=''))

  # take summary values for first 6 params (mu, betaT, sigmas x4) and last param (lp__)
  fitSumm <- summary(stanFit)$summary[c(1:6,97), mInd]

  # compute posterior log-ratio: log(sigmaP/sigmaT), take integral>0, mean, median
  posterior <- as.matrix(stanFit,pars=c("sigma_p","sigma_t"))
  logPTratio <- log(posterior[,1]/posterior[,2])
  CpGscore <- c(mean(logPTratio > 0), mean(logPTratio), median(logPTratio))
  fitSumm <- rbind(fitSumm,CpGscore)

  # print kable of summary values
  print(kable(fitSumm))
  cat("  \n")

  n_eff_sum <- n_eff_sum + fitSumm[,3]
}
mtp <- proc.time() - ptm
